---
title: "tidywoe"
subtitle: "A tidy way to put WoE version of categorical variables into data.frame."
author: "Athos Petri Damiani"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Installation

Currently is available only on Github. To install it, run:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("athospd/tidywoe")
```

# Using tidywoe 

There is tow functions that matter:

- `add_woe()`
- `woe_dictionary()`

## add_woe()

To simply add woe version of your variables from your sample data, just run the follow:

```{r, warning=FALSE, eval = FALSE, message=FALSE, echo = TRUE}
library(dplyr)
library(tidywoe)

# install.packages("FactoMineR")
data(tea, package = "FactoMineR")

tea %>%
  select(breakfast, how, where, price) %>% 
  add_woe(breakfast) %>%
  head
```


```{r, warning=FALSE, echo = FALSE, eval = TRUE, message=FALSE}
library(dplyr)
library(tidywoe)

# install.packages("FactoMineR")
data(tea, package = "FactoMineR")

tea %>%
  select(breakfast, how, where, price) %>% 
  add_woe(breakfast) %>%
  head %>%
  knitr::kable()
```

## woe_dictionary()

The function `add_woe()` uses a dictionary to map the categories with the respective woe value of each explanatory variable.

```{r, message=FALSE, warning=FALSE, echo = TRUE, eval=FALSE}
tea %>%
  select(breakfast, how, where, price) %>% 
  woe_dictionary(breakfast) %>%
  head
```


```{r, message=FALSE, warning=FALSE, echo = FALSE, eval=TRUE}
tea %>%
  select(breakfast, how, where, price) %>% 
  woe_dictionary(breakfast) %>%
  head %>%
  knitr::kable()
```

## Using a custom woe_dictionary

One might want to tweak some woe values to fix the order of the effects of a given variable. That's the why `add_woe()` can receive a custom dictionary through the `.woe_dictionary` argument.

The easy way to do this is to make a previous dictionary with the `woe_dictionary()` as the start point and tweak it to achieve the fixes needed. See below:

```{r, warning=FALSE, message=FALSE, echo = TRUE, eval=FALSE}
# biuld a initial dictionary
tea_woe_dic <- tea %>%
  select(breakfast, how, where, price) %>% 
  woe_dictionary(breakfast)

# tweak it a little bit
tea_woe_dic_tweaked <- tea_woe_dic %>% mutate(woe = if_else(variable == "price" & explanatory == "p_unknown", 0, woe))

# pass it as an argument of add_woe()
tea %>%
  select(breakfast, how, where, price) %>% 
  add_woe(breakfast, .woe_dictionary = tea_woe_dic_tweaked) %>%
  head
```


```{r, warning=FALSE, message=FALSE, echo = FALSE, eval=TRUE}
# biuld a initial dictionary
tea_woe_dic <- tea %>%
  select(breakfast, how, where, price) %>% 
  woe_dictionary(breakfast)

# tweak it a little bit
tea_woe_dic_tweaked <- tea_woe_dic %>% mutate(woe = if_else(variable == "price" & explanatory == "p_unknown", 0, woe))

# pass it as an argument of add_woe()
tea %>%
  select(breakfast, how, where, price) %>% 
  add_woe(breakfast, .woe_dictionary = tea_woe_dic_tweaked) %>%
  head %>%
  knitr::kable()
```
